waRne - Putting cricket analytics in a spin

### 2020-10-02 James_paper_copy.Rmd
rpo plots

bowler_vars <- 
  player_database %>% 
  dplyr::select(-c(batting_style, player_name)) %>% 
  drop_na()%>% 
  mutate(pace_spin = bowling_style, "Right-arm fast" = "Pace",
                                "Slow left-arm orthodox"= "Spin",
                                "Legbreak"= "Spin",
                                "Right-arm medium"= "Pace",
                                "Left-arm medium-fast"= "Pace",
                                "Right-arm offbreak"= "Spin",
                                "Legbreak googly"= "Spin",
                                "Left-arm fast"= "Pace",
                                "Right-arm medium-fast"= "Pace",
                                "Left-arm fast-medium"= "Pace",
                                "Right-arm fast-medium"= "Pace",
                                "Left-arm medium"= "Pace",
                                "Slow left-arm chinaman"= "Spin",
                                "Right-arm offbreak, Legbreak"= "Spin",
                                "Right-arm medium, Right-arm offbreak"= "Spin",
                                "Left-arm medium, Slow left-arm orthodox, 
                                               Slow left-arm chinaman"= "Spin", 
                                "Left-arm medium, Slow left-arm orthodox, Slow left-arm chinaman" = "Spin")

@@

```{r}
# graph of powerplay

@@

# Dot plot for openers
cricjam_data_big %>% 
  dplyr::filter(inning == 2, 
                batsman_id %in% opening_batsman_of_interest, 
                # batsman_balls_faced_this_match<16, 
                !is.infinite(required_run_rate), 
                over<6) %>% 
  group_by(batsman_name) %>% 
  summarise(change = mean(change_in_requireed_run_rate), 
            total_balls_faced = n()) %>% 
  ggplot(aes(x = total_balls_faced, y = change, label = batsman_name)) +
  geom_point(size = 2) +
  xlim(c(50, 1250))+
  geom_text(vjust = 1.4, color = "black") +
  theme_minimal()+
  labs(x = "Balls Faced", 
       y = "Average Change in Required Run Rate", 
       title = "Top Opening Batsmen T20")
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 16),
        plot.title = element_text(size = 16))

@@

```{r}
# middle overs (enter in the middle)

@@

cricjam_data_big %>%
	dplyr::filter(batsman_id == 28081, 
								batsman_balls_faced_this_match == 1) %>%
	group_by(bucketed_over) %>% 
	summarise(total = n()) %>% 
	ungroup() %>% 
	mutate(mean(total))

@@

@@

pollard %>% 
	bind_rows(maxwell, dhoni, russell) %>% 
	ggplot(aes(x = bucketed_over, y = avg, group = batsman_partner, color = batsman_partner)) +
  geom_smooth()+
  geom_hline(yintercept = 0)+
  labs(y = "Required Run Rate", 
       x = "Over", 
       title = "Change in Required Run Rate While Batting (or Partnering)", 
  		 color = "")+
  facet_wrap(~batsman_name)+
  theme_minimal()+
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 16),
        plot.title = element_text(size = 16))

@@

>>>>

bowler_vars <- 
  player_database %>% 
  dplyr::select(-c(batting_style, player_name)) %>% 
  drop_na()%>% 
  mutate(pace_spin = case_when(bowling_style == "Right-arm fast" ~ "Pace",
                                bowling_style == "Slow left-arm orthodox"~ "Spin",
                                bowling_style == "Legbreak"~ "Spin",
                                bowling_style == "Right-arm medium"~ "Pace",
                                bowling_style == "Left-arm medium-fast"~ "Pace",
                                bowling_style == "Right-arm offbreak"~ "Spin",
                                bowling_style == "Legbreak googly"~"Spin",
                                bowling_style == "Left-arm fast"~ "Pace",
                                bowling_style == "Right-arm medium-fast"~ "Pace",
                                bowling_style == "Left-arm fast-medium"~ "Pace",
                                bowling_style == "Right-arm fast-medium"~ "Pace",
                                bowling_style == "Left-arm medium"~ "Pace",
                                bowling_style == "Slow left-arm chinaman"~ "Spin",
                                bowling_style == "Right-arm offbreak, Legbreak"~ "Spin",
                                bowling_style == "Right-arm medium, Right-arm offbreak"~ "Spin",
                                bowling_style == "Left-arm medium, Slow left-arm orthodox, 
                                               Slow left-arm chinaman"~ "Spin", 
                                bowling_style == "Left-arm medium, Slow left-arm orthodox, Slow left-arm chinaman" ~ "Spin"))

@@

```{r, message=FALSE, warning=FALSE, echo = FALSE, error=FALSE, results=FALSE}
# graph of powerplay

@@

# Dot plot for openers
cricjam_data_big %>% 
  dplyr::filter(inning == 2, 
                batsman_id %in% opening_batsman_of_interest, 
                # batsman_balls_faced_this_match<16, 
                !is.infinite(required_run_rate), 
                over<6) %>% 
  group_by(batsman_name) %>% 
  summarise(change = mean(change_in_requireed_run_rate), 
            total_balls_faced = n()) %>% 
  ggplot(aes(x = total_balls_faced, y = change, label = batsman_name)) +
  geom_point(size = 2) +
  xlim(c(50, 1250))+
  geom_text(vjust = 1.4, color = "black") +
  theme_minimal()+
  labs(x = "Balls Faced", 
       y = "Average Change in Required Run Rate", 
       title = "Top Opening Batsmen T20")

@@

```{r, message=FALSE, warning=FALSE, echo = FALSE, error=FALSE, results=FALSE}
# middle overs (enter in the middle)

@@

```{r, message=FALSE, warning=FALSE, echo = FALSE, error=FALSE, results=FALSE}

# cricjam_data_big %>%
# 	dplyr::filter(batsman_id == 28081, 
# 								batsman_balls_faced_this_match == 1) %>%
# 	group_by(bucketed_over) %>% 
# 	summarise(total = n()) %>% 
# 	ungroup() %>% 
# 	mutate(mean(total))

@@

```{r, message=FALSE, warning=FALSE, echo = FALSE, error=FALSE, results=FALSE}

@@

pollard %>% 
	bind_rows(maxwell, dhoni, russell) %>% 
	ggplot(aes(x = bucketed_over, y = avg, group = batsman_partner, color = batsman_partner)) +
  geom_smooth()+
  geom_hline(yintercept = 0)+
  labs(y = "Required Run Rate", 
       x = "Over", 
       title = "Change in Required Run Rate While Batting (or Partnering)", 
  		 color = "")+
  facet_wrap(~batsman_name)+
  theme_minimal()

@@

## RPO

rpo_players <- c(
	# 230559, #pollard
	# 34102, #rohit sharma
	33335, #suresh raina
	28235, #siihkar dhawan
  # 32498, #yusef pathan
  # 321777, #david miller
  # 44936, #AB De Villiers
  # 276298, #andre russell
  # 8180, #shane watson
  34102, #rohit sharma
  # 51880, #gayle
  # 325012, #stoinis
  # 272450, #MR Marsh
  # 28081, #Dhoni
  253802, #kholi
  # 5334, #finch
  219889 #warner
  # 326434, #carey
  # 267192, #smith
  # 505120, #agar
  # 530011, #head
  # 5961, #henriques
  # 325026 #maxwell
  # 25913, #nabu
  # 4864, #christian,
  # 270484, #faulkner
  # 230193, #wade
  # 233680, #wells
  # 500268, #turner
  # 334337, #handscomb
  # 603410, #mcdermott
  # 308798  #short
  # 787987, #labuschagne
  # 1124282, # josh phillipe
  # 326637, #Chris Lyrnn
  # 5939, #hussey
  # 931581 #pant
  # 308967, # job buttler
  # 249866  # alex hales
)

cricjam_data_big %>% 
	dplyr::filter(batsman_id %in% rpo_players, 
								!is.na(pace_spin)) %>% 
	group_by(pace_spin, batsman_name) %>% 
	mutate(rpo = sum(runs)/n()*6) %>% 
	dplyr::summarise(mean_rpo = mean(rpo)) %>% 
	ggplot(aes(x = mean_rpo, y = batsman_name, color = pace_spin))+
	geom_point() +
	theme_minimal()+
	labs(x = "RPO", 
			 title = "RPO By Bowling Style", 
			 y = "", 
			 color = "")
